---
title: "regmods-006_project scratch work"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Preliminaries

```{r message=FALSE}
library("Hmisc")
library("MASS")
library("car")
library("dplyr")
library("ggplot2")
library("GGally")
```

This is a suite of diagnostics that one can run on model fits, so I might as well package them up in a function.

```{r}
dfit <- function(fit, summary = TRUE, ANOVA = TRUE,
                 outlierTest = TRUE, QQ = TRUE, leveragePlots = TRUE, Cook = TRUE,
                 influencePlot = TRUE,
                 ncvTest = TRUE, residualPlots = TRUE,
                 ResidFitted = TRUE, ScaleLocation = TRUE, ResidLeverage = TRUE,
                 vif = TRUE) {
    ## SUMMARY ##
    if (summary) {
        print("SUMMARY:")
        print(summary(fit))
    }
    if (ANOVA) {
        print("ANOVA:")
        print(anova(fit))
    }
    
    ## OUTLIERS, INFLUENCE, AND LEVERAGE ##
    if (outlierTest) {
        print("OUTLIER TEST:")
        print(outlierTest(fit))
    }
    if (QQ) {
        plot(fit, which = 2)
    }
    if (leveragePlots) {
        print(leveragePlots(fit, main = "Leverage Plots"))
    }
    if (Cook) {
        cutoff <- 4/((nrow(fit$model)-length(fit$coefficients)-2)) # D > 4/(n-k-1)  
        plot(fit, which=4, cook.levels=cutoff) # Cook's Distance
    }
    if (influencePlot) {
        print(influencePlot(fit, main="Influence Plot",
                            sub="Circle size is proportial to Cook's Distance"))
    }

    
    ## RESIDUALS AND HOMOSKEDASTICITY ##
    if (ncvTest) {
        print("NCV TEST:")
        print(ncvTest(fit))
    }
    if (residualPlots) {
        print(residualPlots(fit, main = "Residual Plots:"))
    }
    if (ResidFitted) {
        plot(fit, which = 1)
    }
    if (ScaleLocation) {
        plot(fit, which = 3)
    }
    if (ResidLeverage) {
        plot(fit, which = 5)
    }
    
    ## MULTICOLLINEARITY ##
    if (vif) {
        print("VIF:")
        print(vif(fit))
    }
}
```

Convert factor variables.
```{r}
mtcars2 <- mtcars %>%
    mutate(cyl = factor(cyl),
           vs = factor(vs, levels = c(0, 1), labels = c("V", "S")),
           am = factor(am, levels = c(0, 1), labels = c("Automatic", "Manual")),
           gear = factor(gear, levels = c(3, 4, 5)))
```

## EDA

Basic summary:
```{r}
mpg_by_am <- mtcars2 %>%
    group_by(am) %>%
    summarise(mean(mpg), sd(mpg))
mpg_by_am
ggplot(data = mtcars2, aes(x = am, y = mpg)) +
    geom_boxplot() +
    xlab("Transmission") +
    ylab("MPG")
```

Scatterplot matrices with correlation tables.

(Note. Significant when $|r| > 0.349$.)

```{r cache=TRUE, fig.width=10, fig.height=10}
mtcars_num <- mtcars2 %>%
    select(mpg, disp, hp, drat, wt, qsec, carb)
mtcars_cat <- mtcars2 %>%
    select(mpg, wt, cyl, vs, am, gear)
ggpairs(mtcars_num)
ggpairs(mtcars_cat)
```

(Note that our factor variables are all associated with weight as well.)



## mpg ~ am

(This is just one-way ANOVA with two categories, which is really just a two-sample t-test.)

```{r}
fit1 <- lm(mpg ~ am, data = mtcars2)
dfit(fit1, vif = FALSE)
```

Equivalent to a two-sample t-test with equal variances assumed. ($F = t^{2}$.)

```{r}
fit_t1 <- t.test(mpg ~ am, var.equal = TRUE, data = mtcars2)
fit_t1
all.equal(anova(fit1)$F[1], (fit_t1$statistic)^2, check.attributes = FALSE)
```

It is more correct, however, to perform Welch's t-test:

```{r}
fit_t2 <- t.test(mpg ~ am, data = mtcars2)
fit_t2
```

## mpg ~ .

```{r}
fit_all <- lm(mpg ~ ., data = mtcars2)
dfit(fit_all)
```

## mpg ~ am + qsec

```{r}
fit2 <- lm(mpg ~ am + qsec, data = mtcars2)
dfit(fit2)
```

## mpg ~ am + wt

```{r}
fit3 <- lm(mpg ~ am + wt, data = mtcars2)
dfit(fit3)
```

## mpg ~ am + wt + qsec

Recommended by `stepAIC`.

```{r}
fit4 <- lm(mpg ~ am + wt + qsec, data = mtcars2)
dfit(fit4)
```

Note that we have improvement in model fit for all steps except the last.
```{r}
anova(fit1, fit3, fit4, fit_all)
anova(fit1, fit2, fit4, fit_all)
```